### 2023-06-13 application.tex
up

We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient. The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
\alain{vérifier que la plus grande existe}
Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

@@

\begin{theorem}{}
Resilience is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear sets.
\end{theorem}

@@

\begin{theorem}{}
Resilience is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear sets.
\end{theorem}

\begin{proof}
The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^d$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
% For Z-VAS we observe that we are asking for inclusion between linear sets. 
% Again let us without loss of generality hence let us consider only the case when $\Safe$ is semilinear and $\Bad = S \setminus \Safe$ is semilinear too.
\mathieu{Expliciter davantage la preuve ? $\pred^*(\Safe)$ semilinéaire car définissable en Presburger ? Puis inclusion de semilinéaires décidables.}
\end{proof}

@@

\begin{figure}

\begin{center}
\begin{tabular}{ | l | c | c | c | r |}
\hline   \Safe~\Bad 
		& $\uparrow$~ $\downarrow$~ 
		 & $\downarrow$~ $\uparrow$~ 
 \\ \hline
   RP  & Decidable (Thm~\ref{down-up})   & Decidable (Thm~\ref{RP VASS})
    \\ \hline
   BRP   &  Decidable (Thm~\ref{down-up})  & Decidable (Thm~\ref{brp VASS})
    \\ \hline
      kRP       & Decidable (Thm~\ref{down-up}) & Decidable (Thm~\ref{brp VASS})
       \\ \hline
   sRP    	& Decidable (Thm~\ref{post srp})	 & ??
    \\ \hline
   BsRP     &  Decidable (Thm~\ref{ref ozkan})	 & ?? 
    \\ \hline
      ksRP        & Decidable (Thm~\ref{ref ozkan}) & ?? 
       \\ \hline

\end{tabular}
\end{center}
	\caption{Decidability results for VASS.}
\end{figure}

\begin{proof}
Consider a given $k \in \N$,
 $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed, and
consider a given vector addition system $V$.
Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.

The set $\Bad$ admits a finite basis $B_\Bad$.
Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.

All configurations reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
and thus, are in $\Bad$, by upward-closedness.

Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
\end{proof}

\iffalse
\mathieu{
resilience never hold in the case of VAS. 
What happens when you add states ? \\
So you consider for instance $q_0(u_k)$ instead of just $u_k$. \\
Everything reachable from $q_0(u_k)$ is still of the form $q(u)$ with u greater than $v_{\Bad}$
(we'd have $v_{\Bad}$ sum of the coordinate of the elements from the basis of $\Bad$ when not considering states).
So now, let us fix some state $q$.
Either there is an element of the basis of $\Bad$ with state $q$.
Then $q(u)$ is necessarily in $\Bad$ (If this hold for all $q \in Q$, then again $k$-resilience does not hold).
Either there is none.
If there is none, then $q(\omega)$ is in the complement of $\Bad$, i.e. $\Safe$.
Then, of course there are probably other elements in $\Safe$ but from
the perspective of $\Bad$ and the extremely high elements $q_i(u_k)$  these ‘smaller’ elements are all but unreachable. 
And the question becomes rather `can we reach $q(\omega)$ from every element of $\Bad$ save from maybe the few smaller ones ?’
Since $q(\omega)$ here is upward-closed this is decidable i.e. it is asking coverability questions. 
It’s not 100\% the question «asking, for every element in the basis of $\Bad$, whether or not $q(0)$ is coverable from it», since you could have, for smaller elements in $\Bad$, other elements of $\Safe$ you can reach, but it seem close.
This would lead to a decision scheme.
Need to write this more formally/properly.}
\fi

@@

\begin{proof}
\mathieu{Donner l'idée et se référencer à l'appendix~\ref{appendix brp VASS} pour la preuve complête.}
\end{proof}

>>>>

We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient. The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
%\alain{vérifier que la plus grande existe}
Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

@@

\begin{theorem}{}
Resilience is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear.
\end{theorem}

@@

\begin{theorem}{}
Resilience is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear.
\end{theorem}

\begin{proof}
The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^d$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
% For Z-VAS we observe that we are asking for inclusion between linear sets. 
% Again let us without loss of generality hence let us consider only the case when $\Safe$ is semilinear and $\Bad = S \setminus \Safe$ is semilinear too.
% \mathieu{Expliciter davantage la preuve ? $\pred^*(\Safe)$ semilinéaire car définissable en Presburger ? Puis inclusion de semilinéaires décidables.}
\end{proof}

@@

\iffalse
\begin{figure}
\begin{center}
\begin{tabular}{ | l | c | c | c | r |}
\hline   \Safe~\Bad 
		& $\uparrow$~ $\downarrow$~ 
		 & $\downarrow$~ $\uparrow$~ 
 \\ \hline
   RP  & Decidable (Thm~\ref{down-up})   & Decidable (Thm~\ref{RP VASS})
    \\ \hline
   BRP   &  Decidable (Thm~\ref{down-up})  & Decidable (Thm~\ref{brp VASS})
    \\ \hline
      kRP       & Decidable (Thm~\ref{down-up}) & Decidable (Thm~\ref{brp VASS})
       \\ \hline
   sRP    	& Decidable (Thm~\ref{post srp})	 & ??
    \\ \hline
   BsRP     &  Decidable (Thm~\ref{ref ozkan})	 & ?? 
    \\ \hline
      ksRP        & Decidable (Thm~\ref{ref ozkan}) & ?? 
       \\ \hline

\end{tabular}
\end{center}
	\caption{Decidability results for VASS.}
\end{figure}
\fi

\begin{proposition}
{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
\end{proposition}

\begin{proof}
Consider a given $k \in \N$,
 $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed, and
consider a given vector addition system $V$.
Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
The set $\Bad$ admits a finite basis $B_\Bad$.
Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.

All configurations reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
and thus, are in $\Bad$, by upward-closedness.
Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
\end{proof}

This all changes when you add control states. For a VASS with control states in $Q$, a set $\Bad = \uparrow \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none. If there is none, then 
$\uparrow q(\textbf{0})$
 is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
 $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$. This lead to a decision procedure, written in detail in Appendix~\ref{appendix brp VASS}, which lead to the following decidability result:

@@

% \begin{proof}
% \mathieu{Donner l'idée et se référencer à l' pour la preuve complête.}
% \end{proof}
